\TheTitle

### 2019-07-26 Paper1/framework.tex
introduction and framework nicer

A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}. This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}. By simply taking moments of a given deterministic flux, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available. Therefore, we call the proposed numerical method \textit{semi-intrusive}.

Moreover, by computing the numerical flux with a quadrature rule, we minimize numerical costs: When physical fluxes or source terms in the deterministic equation are polynomials, numerical fluxes for the SG moment system \eqref{eq:SGMomentSystem} can be computed analytically before running the program. Even when terms in the deterministic problem are not polynomials, an approximation which again allows a precomputation can be constructed \cite{debusschere2004numerical}. See for example \cite{hu2015stochastic,tryoen2010instrusive} for the numerical flux computation or \cite{hu2016stochastic} for the computation of source terms. However, analytically precomputing terms in the SG moment system is often less efficient than using techniques as stratified or latin hypercube sampling \cite[Section 6]{ghanem1998stochastic}, see Appendix~\ref{app:costNumFlux}. When not using a quadratic entropy in the IPM method, integrals cannot be evaluated analytically and quadrature rules are not exact. We will study the approximation properties of IPM with different quadrature orders in Section~TODO

>>>>

A straight-forward implementation is ensured by the choice of the numerical flux \eqref{eq:numericalFluxIPM}. This choice of the numerical flux is common in the field of transport theory, where it is called the \textit{kinetic flux}. By simply taking moments of a given numerical flux for the deterministic problem, the method can easily be applied to various physical problems whenever an implementation of $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available. Therefore, we call the proposed numerical method \textit{semi-intrusive}.

Intrusive numerical methods which compute arising integrals analytically and therefore directly depend on the moments (i.e. they do not necessitate the evaluation of the gPC expansion on quadrature points) can be constructed by performing a gPC expansion on the system flux directly \cite{debusschere2004numerical}. Examples can be found in \cite{hu2015stochastic,hu2016stochastic,tryoen2010instrusive}. While the analytic computation of arising integrals is not always more efficient \cite[Section 6]{ghanem1998stochastic}, it can also complicate recycling a deterministic solver. See \ref{app:costNumFlux} for a comparison of numerical costs when using Burgers' equation. However, when not using a quadratic entropy in the IPM method or when the physical flux of the deterministic problem is not a polynomial, it is not clear how many quadrature points the numerical quadrature rule requires to guarantee a sufficiently small quadrature error. We will study the approximation properties of IPM with different quadrature orders in Section~TODO

### 2019-07-26 Paper1/introduction.tex
introduction and framework nicer

The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem. Our numerical studies show the following main results:
\begin{itemize}
\item The standard IPM method cannot compete with SC and SG. However when using the discussed acceleration techniques, IPM reaches a given accuracy with the shortest run time.
\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM leads to better variance approximations.
\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods, even for one spatial dimension to reach the same accuracy level.
\end{itemize}
Furthermore, we present a semi-intrusive method, which facilitates the task of implementing general intrusive methods. The method only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.

>>>>

The effectiveness of these acceleration ideas are tested by comparing results with stochastic-Collocation for the uncertain NACA test case as well as a bent shock tube problem. Our numerical studies show the following main results:
\begin{itemize}
\item In our test cases, the need to solve the dual problem \eqref{eq:dualProblem} when using the IPM method leads to a significantly higher run time than SC and SG. However when using the discussed acceleration techniques, IPM requires the shortest time to reach a given accuracy.
\item Comparing SG with IPM, one observes that for the same number of unknowns, SG yields more accurate expectation values, whereas IPM shows improved variance approximations.
\item By studying aliasing effects, we show that SC requires a higher number of unknowns than intrusive methods (even for one spatial dimension) to reach the same accuracy level.
\end{itemize}
The IPM and SG calculations use a semi-intrusive numerical method, since the discretization allows recycling a given deterministic code to generate the IPM solver. While facilitating the task of implementing general intrusive methods, this framework allows a fair comparison of intrusive and non-intrusive methods, since the intrusive method can be based on the same deterministic code as the code used in a black-box fashion for stochastic-Collocation.
